\large{\bf Looking forwards and backwards through locally regulated populations }

### 2022-04-24 forwards-and-backwards.tex
errors squashed!

%%%%%%% notation
\newcommand{\DG}{\mathcal{B}}  % generator of the dispersal process
\newcommand{\meanq}{\vec b}    % mean of dispersal, times theta
\newcommand{\covq}{\Sigma}     % covariance matrix of dispersal, times theta
\newcommand{\kernel}{\rho}  % interaction kernels
\newcommand{\smooth}[1]{\kernel_{#1} \! * \!}  % convolution by the interaction kernel
\newcommand{\wavespeed}{\mathfrak{c}}    % speed (vector) of a wave
\newcommand{\Lgen}{\mathcal{L}}    % generator of a lineage
\newcommand{\Pgen}{\mathcal{P}}    % generator of the population process
\newcommand{\lp}{\xi}              % process with levels
\newcommand{\labelspace}{\mathcal{I}} % space of labels

@@

@@

\begin{definition}[Dispersal generator]
    \label{def:dispersal_generator}
    Recall that we have defined the dispersal kernel,
    $q_\theta(x, dy)$,
    to be the density of a multivariate Normal
    with mean $\meanq/\theta$ and covariance matrix $\covq/\theta$.
    This implies that if we define, for $f : \IR^d \to \IR$,
    \begin{align}
    \DG f(x) = \sum_i \meanq_i \partial_{x_i} f(x)
        + \sum_{ij} \covq_{ij} \partial_{x_i} \partial_{x_j} f(x)
    \end{align}
    then
    \begin{align}
        \theta \int \left(
            f(y) - f(x)
        \right) q_\theta(x, dy)
    \to \DG f(x) 
        \qquad \text{as } \theta \to \infty .
    \end{align}
\end{definition}

@@

\begin{corollary}[Deterministic limits]
    \label{cor:nonlocal_pde_limits}
    \comment{(TODO: remove "has a density" and rephrase appropriately)}
    In the context of Theorem~\ref{thm:nonlocal_convergence},
    if the limit is deterministic (i.e., $\theta / N \to \alpha = 0$), $\eta_0$ has a density,
    and \comment{additional conditions? can we say something like this??}
    then any limit $\eta_t(dx)$ has a density $\varphi(t, x)$,
    which is a weak solution to the nonlocal partial differential equation
    \begin{equation} \label{eqn:nonlocal_pde}
        \partial_t \varphi(t, x)
        =
        r\left(x, \smooth{r} \varphi(t,x) \right)
        \mathcal{B} \left[
            \varphi(t, \cdot)
            \gamma\big( x, \smooth{\gamma} \varphi(t, \cdot) \big)
        \right](x)
        +
        \varphi(t, x)
        F\left(x, \smooth{F} \varphi(t,x) \right)
        .
    \end{equation}
    If equation \eqref{eqn:nonlocal_pde} has a unique solution,
    then $(\eta^\theta_t)_{t \ge 0}$ converges to that solution.

@@

Informally, equation \eqref{eqn:nonlocal_pde} is the PDE
\begin{align} \label{eqn:pde}
    \dot \varphi = r \DG\left( \gamma \varphi \right) + \varphi F ,
\end{align}
where $\DG$ is an elliptic second order differential operator,
and the coefficients $r$, $\gamma$, and $F$
at $x$ depend nonlocally on $\varphi$
(i.e., through a smoothed version of $\varphi$).

@@

say $\eta_t(dx) = \varphi(x) dx$,

@@

    \varphi(x) \gamma(x, \eta_t) r(y, \eta_t)

@@

    \varphi(y)

@@

    \to

@@

    + \sum_{ij} \covq_{ij} \partial_{x_i x_j} f(y) ,
which is just $\DG$ but with the first derivative terms multiplied by -1.
\comment{The $\widetilde \DG$ and $\covq \grad$ notation here is terrible; suggestions?}

@@

@@

    - g(y) \sum_i \meanq_i \partial_{x_i} f(y)
    + g(y) \sum_{ij} \covq_{ij} \partial_{x_i x_j} f(y)
    + \sum_{ij} \partial_{x_i} g(y) \covq_{ij} \partial_{x_j} f(y)  \\

@@

    g(y) \widetilde \DG f(y)
    + (\covq \grad g) \cdot \grad f ,

@@

if we define $\widetilde \DG$ to be the same as $\DG$ but with $-\meanq$ substituted for $\meanq$,
and $(\covq \grad g(y))_i = \sum_j \covq_{ij} \partial_{x_j} g(y)$.

@@

@@

        \varphi_{T-s} \gamma \widetilde \DG f
        + \left(\covq \grad (\varphi_{T-s} \gamma) \right) \cdot \grad f 

@@

        \widetilde \DG f
        + \left(\covq \grad \log(\varphi_{T-s} \gamma) \right) \cdot \grad f 

@@

of the production of juveniles, $\grad \log(\phi_{T-s} \gamma)$

@@

Potential reproduction events occur at times $\tau_a + s$
for each point $(s, v, y, \kappa) \in \Pi_a$.
We say `potential' since if the resulting level is greater than $\lambda$,
the event does not happen.
If this is the $j^\text{th}$ point in $\Pi_a$,
the new label is $aj$, the (potential) birth time is $\tau_{aj} = \tau_a + s$.
For convenience define
$$
    \ell(x, y, \eta, v)
    =
    \frac{
        v
    }{
        2 \lambda^{-1} \theta \gamma(x, \eta) r(x + y, \eta)
    } ,
$$
and using this, the (potential) new level is
\begin{equation*}
    U_{aj}(\tau_{aj})
    =
    U_a(\tau_{aj})
    +
    \ell(X_a(\tau_{aj}-), y, \eta(\tau_{aj}-), v) .
\end{equation*}
If $U_{aj}(\tau_{aj}) < \lambda$,
the new individual labeled $aj$ is produced,
and $\kappa$ determines which label is associated with the new location,
so
\begin{align*}
    X_{aj}(\tau_a + s)
    &=
    X_a(\tau_a+s-) + (1 - \kappa) y .
\end{align*}
On the other hand
if $U_{aj}(\tau_a+s) \ge \lambda$, then $X_a$ is unchanged and $X_{aj}$ is undefined,
so
\begin{align} \label{eqn:X_line_of_descent}
    X_a(\tau_a+s)
    &=
    X_a(\tau_a+s-) + \kappa y \mathbbm{1}_{U_{aj}(\tau_{aj}) < \lambda} .
\end{align}
Recall that the parental \emph{individual} always retains their spatial location,
so that $\kappa = 0$ corresponds to the parent being assigned a new level,
and our line of descent switching to the offspring.

@@

First, for each $1 \le k < n$
let $(s_k, v_k, y_k, \kappa_k)$ be the $a_{k+1}^\text{th}$ point in $\Pi_{(a)_k}$
(this is the sequence of jumps taken by the line of descent).
Note that $s_k = \tau_{(a)_k}$, and also set $s_0 = 0$.
Let $\Gamma_a$ by the point process ``experienced by the line of descent leading to $a$'',
which is for $A \subset [0,\infty)^2 \times \IR^d \times \{0,1\}$
\comment{This is wrong, because $s$ is relative to birth time.}
\begin{align*}
\Gamma_a(A)
    &=
    \sum_{k=1}^n
    \Pi_{(a)_k}\left(A \cap [s_{k-1}, s_k) \times [0, \infty) \times \IR^d \times \{0,1\}\right) 
    \\ &\qquad {}
    + \sum_{k=1}^{n-1}
      \delta_{(s_k, v_k, y_k, 1 - \kappa_k)}(A) .
\end{align*}
Note that we have included the points at which the line of descent's level jumps,
flipping $\kappa$ at these points,
and as a result $\Gamma_a$ has the same distribution as $\Pi_a$
restricted to $[0, \sigma_a)$.
Then for $0 \le t \le \sigma_a$,
\begin{align*}
    X_a(t)
    &=
    X_{a_1}(0)
    +
    \int_{
        [0, t) \times [0, \infty) \times \IR^d \times \{0, 1\}
    }
    \kappa y
    \mathbbm{1}_{
        U_{(a)_{\zeta(s)}}(s) + \ell( X_{(a)_{\zeta(s)}}(s-), y, \eta(s-), v) < \lambda
    }
    \Gamma_{a}(ds, dv, dy, d\kappa) 
\end{align*}
and
\begin{align*}
    U_a(t)
    &=
    U_{a_1}(0)
    +
    \sum_{k=1}^{\zeta_a(t) - 1}
    \frac{
        v_k
    }{
        2 \lambda^{-1} \theta \gamma(X_{(a)_k}(s_k-), \eta(s_k-))
        r(X_{(a)_k}(s_k-) + y_k, \eta(s_k-))
    }
    \\ & \qquad {}
    +
    \int_0^t
    \bigg(
        \frac{\theta}{\lambda} \gamma(X_a(s),\eta(s))
        \int_{\IR^d} r(z,\eta(s)) q(X_a(s),dz) U_a(s)^2
    \\ & \qquad\qquad  \qquad {}
        -
        b(X_a(s),\eta(s)) U_a(s)
    \bigg)
    dt .
\end{align*}

%% %% %% %% %% %% %% %% %% %% %% %%
\subsection{Limiting processes for lines of descent}

@@

Supposing that they do,
consider how a line of descent $(X_a(t), U_a(t))$ evolves.
The line of descent throws off a new line of descent at a higher level
when there is a point $(s, v, y, \kappa)$ in $\Pi_a$ with 
\begin{equation} \label{eqn:line_points}
v < 2 (\lambda - U_a(t)) \lambda^{-1} \theta \gamma(X_a(t), \eta) r(X_a(t)+y, \eta) .
\end{equation}
Since the mean measure of the $v$ coordinate is Lebesgue,
$\theta/\lambda \to \delta$,
and $q(x, dy) \to \delta_x(dy)$,
this corresponds in the limit to new lines of descent being thrown off at all times,
i.e., according to a Poisson process with rate
$$
\delta \gamma(X_a(t), \eta) r(X_a(t), \eta) 
$$
and uniform intensity on higher levels.
Again, since in the limit the dispersal distance is zero,
new levels are produced at the same spatial location.
However, this affects the location of the line of descent:
for each new line of descent, there is a $1/2$ chance
that the line of descent switches location (i.e., moves to $X_a(t) + y$).
Taking $g$ to be a suitable test function on $\IR^d$, 
and rewriting~\eqref{eqn:line_points},
the generator of this motion is, when the level is $u$,
\begin{align*}
    g(x)
    &\mapsto
    \left(1 - \frac{u}{\lambda}\right) \gamma(x, \eta)
    \theta \int_{\IR^d} r(x+y, \eta) (g(x+y) - g(x)) q(x, dy) \\
    &=
    \left(1 - \frac{u}{\lambda}\right) \gamma(x, \eta)
    \bigg\{
        \theta \int_{\IR^d} (r(x+y, \eta) g(x+y) - r(x, \eta) g(x)) q(x, dy) \\
        &\qquad \qquad {}
        -
        \theta \int_{\IR^d} (r(x+y, \eta) - r(x, \eta) ) g(x) q(x, dy) 
    \bigg\} \\
    &\to
    \gamma(x, \eta)
    \left(
        \DG(rg)(x) - g(x) \DG(r)(x)
    \right) .
\end{align*}
Notice that the factors of 2 have canceled,
and that the result is independent of $u$.

@@

To write out the joint motion of spatial coordinate and levels,
it is useful to work out the differential operator above in more detail.
Recall that $\DG g(x) = \sum_i \meanq_i \partial_i g(x) + \sum_{ij} \covq_{ij} \partial_{ij} g(x)$,
and for the moment write $r(x)$ for $r(\cdot, \eta(\cdot))(x)$,
so that
\begin{align*}
\DG(rg)(x) - g(x) \DG(r)(x)
    &=
    r(x) \sum_i \meanq_i \partial_i g(x)
    + 2 \sum_{ij} \meanq_i \partial_i r(x) \covq_{ij} \partial_j g(x)
    + r(x) \sum_{ij} \covq_{ij} \partial_{ij} g(x) \\
    &=
    r(x) \left\{
        \left(
        \meanq
        + 2 \covq \grad \log r(x)
        \right)
        \cdot
        \grad g(x)
        +
        \sum_{ij} \covq_{ij} \partial_{ij} g(x)
    \right\} .
\end{align*}

@@

Now, let $\beta(x, \eta) = \meanq + 2 \Sigma \grad \log r(x, \eta)$,
and associate with each label an independent $d$-dimensional Brownian motion $W_a$,
and let $\covq = K K^T$.
Then, we can summarize the results above by saying that
in the infinite density limit,
the spatial locations and level of each line of descent $a$
behave for $\tau_a \le t < \sigma_a$ as
\begin{align*}
X_a(t)
    &=
    X_a(\tau_a)
    + \int_{\tau_a}^{t}
    \bigg(
        \beta(x, \eta(s)) ds
        +
        K dW_a(s)
    \bigg)
    ds  \\
U_a(t)
    &=
    U_a(\tau_a)
    + \int_{\tau_a}^{t}
    \bigg(
        \delta \gamma(X_a(s),\eta(s))
        r(x,\eta(s)) U_a(s)^2
\\ &\qquad \qquad \qquad {}   
        -
        \left\{
            \gamma(X_a(s),\eta(s)) \DG r(X_a(s),\eta(s))
            + F(X_a(s), \eta(s))
        \right\}
        U_a(s)
    \bigg)
    ds .
\end{align*}
To complete the description,
for each label $a \in \labelspace$ let $R_a$ denote an independent Poisson process on $[0, \infty)^2$
with Lebesgue mean measure,
and as before order the points of each in some way.
If the population state at time 0 is $\eta(0)$,
then let $(x_i, u_i)$ be the points of a Poisson process
with mean measure on $\IR^d \times [0, \infty)$ a product of $\eta(0)$ and Lebesgue measure.
\comment{Here I've switched $t$ to be the absolute birth time
rather than relative to the birth time of the parent; I think it's clearer?
It, should be made to agree with the previous section.}
For each point begin a line of descent
with label $i$, location $X_i(0) = x_i$, level $U_a(0) = u_i$, and birth time $\tau_i = 0$.
Each point in the $R_a$ denotes a potential birth time for $a$:
if the $j^\text{th}$ point in $R_a$ is $(t, v)$,
$\tau_a \ge t$, and and $U_a(t) < \infty$,
then a new line of descent with label $aj$ is produced,
with birth time $\tau_{aj} = t$, location $X_{aj}(t) = X_a(t)$, and level
$$
    U_{aj}(t) = U_a(t)
    + \frac{v}{ 2 \delta \gamma(X_a(t), \eta(t)) r(X_a(t), \eta(t)) } .
$$
\comment{Perhaps the conclusion of the Propostion would be something like:}
We can now express our population processes in the limit using these levels:
\begin{align*}
    \eta_t = \sum_{a : \tau_a > t \;\&\; U_a(t) < \lambda} \delta_{X_a(t)}
    \qquad \text{and} \qquad
    \lp = \sum_{a : \tau_a > t \;\&\; U_a(t) < \lambda} \delta_{(X_a(t), U_a(t))} .
\end{align*}

@@

level. Consequently, for $t \geq \tau$
\begin{align}\label{eq: Spatial evolution of individual lineage}
X_{\tau}(\tau+t)=&X_{\tau}(t)\\
&+\int_{(0,t]\times[0,\infty]\times[0,1]\times\{0,1\}}\mathbbm{1}_{U_{\tau}(\tau+s)+ \frac{v}{2\lambda^{-1} \theta \gamma^{\mathfrak{m}}_{\theta}(X_{\tau}(\tau+s-), \eta(\tau+s-))}<\lambda\}}\kappa    H(X_{\tau}(\tau+s-), \eta(\tau+s),\omega) \\
&\qquad \times \xi_{\tau}(ds,dv,dw,d\kappa).
\end{align}
Starting with an initial particle at level $U_{(0,i)}$ , we can identify a potential line of descent by
taking $(s_1 , v_1 , w_1 , \kappa_1 ) \in \xi_{(0,i)}$ , setting $\tau_1=s_1$ , and then recursively taking $(s_k , v_k , w_k , \kappa_k ) \in \xi_{\tau_k }$ and defining $\tau_{k+1} = \tau_{k} + s_{k+1}$. Each time $\tau$ at which a new level appears has an “ancestory” of this type.\\
We now explain the cases where $\kappa_k=0$ and $\kappa_k=1$. When $\kappa_k=0$, the parent retains the level, so $X_{\tau_{k}}$ remains the same at time $\tau_{k+1}$. When $\kappa_k=1$, the offspring gets its parent's level while the parent inherits a new level sampled between $[U_{\tau_k}(\tau_{k+1}-),\lambda]$. As a result, the spatial location $X_{\tau_{k}}$ jumps from the parent's location to the offspring's location at time $\tau_{k+1}$.  Meanwhile, between $[\tau_k, \tau_{k+1}]$ there maybe further birth events generated by $\xi_{\tau_k}$. If $(s,v,w,\kappa) \in \xi_{\tau_k}$, Equation \eqref{eq: Spatial evolution of individual lineage} holds and gives the spatial evolution of $X_{\tau_k}$ for $t \in [\tau_k, \tau_{k+1}]$.\\

>>>>

%%%%%%% notation
\newcommand{\DG}{\mathcal{B}}  % generator of the dispersal process
\newcommand{\DD}{\mathcal{D}}  % the second-order part of the generator of dispersal
\newcommand{\meanq}{\vec b}    % mean of dispersal, times theta
\newcommand{\covq}{C}     % covariance matrix of dispersal, times theta
\newcommand{\kernel}{\rho}  % interaction kernels
\newcommand{\smooth}[1]{\kernel_{#1} \! * \!}  % convolution by the interaction kernel
\newcommand{\wavespeed}{\mathfrak{c}}    % speed (vector) of a wave
\newcommand{\Lgen}{\mathcal{L}}    % generator of a lineage
\newcommand{\Pgen}{\mathcal{P}}    % generator of the population process
\newcommand{\lp}{\xi}              % process with levels
\newcommand{\labelspace}{\mathcal{I}} % space of labels

@@

$$\lim \frac{\theta}{N} =: \alpha$$

@@

\comment{TODO: $\meanq$ can vary in space, make that clear?}

\begin{definition}[Dispersal generator]
    \label{def:dispersal_generator}
    Recall that we have defined the dispersal kernel,
    $q_\theta(x, dy)$,
    to be the density of a multivariate Normal
    with mean $\meanq/\theta$ and covariance matrix $\covq/\theta$.
    This implies that if we define, for $f : \IR^d \to \IR$,
    \begin{align}
    \DG f(x)
        =
        \sum_{ij} \covq_{ij} \partial_{x_i} \partial_{x_j} f(x)
        + \sum_i \meanq_i \partial_{x_i} f(x)
    \end{align}
    then
    \begin{align}
        \theta \int \left(
            f(y) - f(x)
        \right) q_\theta(x, dy)
    \to \DG f(x) 
        \qquad \text{as } \theta \to \infty .
    \end{align}
    We will also denote the adjoint of $\DG$ by
    $$
    \DG^* f(x)
        =
        \sum_{ij} \covq_{ij} \partial_{x_i} \partial_{x_j} f(x)
        + \sum_i \partial_{x_i} (f(x) \meanq_i) .
    $$
\end{definition}

@@

\begin{corollary}[Deterministic limits]
    \label{cor:nonlocal_pde_limits}
    \comment{(TODO: remove "has a density" and rephrase appropriately)}
    In the context of Theorem~\ref{thm:nonlocal_convergence},
    if the limit is deterministic (i.e., $\theta / N \to \alpha = 0$), $\eta_0$ has a density,
    and \comment{additional conditions? can we say something like this??}
    then any limit $\eta_t(dx)$ has a density $\varphi(t, x)$,
    which is a weak solution to the nonlocal partial differential equation
    \begin{equation} \label{eqn:nonlocal_pde}
        \partial_t \varphi(t, x)
        =
        r\left(x, \smooth{r} \varphi(t,x) \right)
        \DG^* \left[
            \varphi(t, \cdot)
            \gamma\big( x, \smooth{\gamma} \varphi(t, \cdot) \big)
        \right](x)
        +
        \varphi(t, x)
        F\left(x, \smooth{F} \varphi(t,x) \right)
        .
    \end{equation}
    If equation \eqref{eqn:nonlocal_pde} has a unique solution,
    then $(\eta^\theta_t)_{t \ge 0}$ converges to that solution.

@@

Informally, equation \eqref{eqn:nonlocal_pde} is the PDE
\begin{align} \label{eqn:pde}
    \dot \varphi = r \DG^* \left( \gamma \varphi \right) + \varphi F ,
\end{align}
where $\DG$ is an elliptic second order differential operator,
and the coefficients $r$, $\gamma$, and $F$
at $x$ depend nonlocally on $\varphi$
(i.e., through a smoothed version of $\varphi$).
To see why it is $\DG^*$ and not $\DG$ that appears in this equation,
consider the case where dispersal has a mean drift to the right,
so that the drift term in $\DG$ is $d/dx$.
Then, the population density at a location with positive slope
will tend to \emph{decrease}, i.e., proportional to $-d/dx$,
because the density at a location is determined by the offspring that disperse \emph{to}
that location, not away from it.
Furthermore, if dispersal tends to move offspring away from a region
(i.e., if divergence $\grad \cdot \meanq > 0$)
this will tend to decrease the local density.

@@

say $\eta_t(dx) = \varphi_t(x) dx$,

@@

    \gamma(x, \eta_t) r(y, \eta_t) \varphi_t(x) 

@@

    \varphi_t(y)

@@

    &\to

@@

    + \sum_{ij} \covq_{ij} \partial_{x_i x_j} f(y) \\
    = \DD - \meanq \cdot \grad f ,
where $\DD = \sum_{ij} \covq_{ij} \partial_{ij}$.
(Note that this differs from $\DG^*$ if $\meanq$ can vary across space.)

@@

    \theta \int \left\{
        (f(x) g(x) - f(y) g(y)) - f(y) (g(x) - g(y))
    \right\} \widetilde q_\theta(y, dx) \\

@@

        \DD(fg)(x) - f(x) \DD g(x) - \meanq \cdot \left( \grad(fg)(x) - f(x) \grad g(x) \right) \\

@@

        g(x) \DD f(x) + 2 (\covq \grad g(x)) \cdot \grad f(x) - g(x) \meanq \cdot \grad f(x) .

@@

@@

\begin{align} \label{eqn:heuristic_lineage_generator}

@@

        \varphi_{T-s} \gamma (\DD f - \meanq \cdot \grad f)
        + 2 \left(\covq \grad (\varphi_{T-s} \gamma) \right) \cdot \grad f 

@@

        \DD f - \meanq \cdot \grad f
        + 2 \left(\covq \grad \log(\varphi_{T-s} \gamma) \right) \cdot \grad f 

@@

of the production of juveniles, $2 \grad \log(\phi_{T-s} \gamma)$

@@

Potential reproduction events occur at times $\tau_a + s$
for each point $(s, v, y, \kappa) \in \Pi_a$.
We say `potential' since if the resulting level is greater than $\lambda$,
the event does not happen.
If this is the $j^\text{th}$ point in $\Pi_a$,
the new label is $aj$, the (potential) birth time is $\tau_{aj} = \tau_a + s$.
For convenience define
$$
    \ell(x, y, \eta, v)
    =
    \frac{
        v
    }{
        2 \lambda^{-1} \theta \gamma(x, \eta) r(x + y, \eta)
    } ,
$$
and using this, the (potential) new level is
\begin{equation*}
    U_{aj}(\tau_{aj})
    =
    U_a(\tau_{aj})
    +
    \ell(X_a(\tau_{aj}-), y, \eta(\tau_{aj}-), v) .
\end{equation*}
If $U_{aj}(\tau_{aj}) < \lambda$,
the new individual labeled $aj$ is produced,
and $\kappa$ determines which label is associated with the new location,
so
\begin{align*}
    X_{aj}(\tau_a + s)
    &=
    X_a(\tau_a+s-) + (1 - \kappa) y .
\end{align*}
On the other hand
if $U_{aj}(\tau_a+s) \ge \lambda$, then $X_a$ is unchanged and $X_{aj}$ is undefined,
so
\begin{align} \label{eqn:X_line_of_descent}
    X_a(\tau_a+s)
    &=
    X_a(\tau_a+s-) + \kappa y \ind_{U_{aj}(\tau_{aj}) < \lambda} .
\end{align}
Recall that the parental \emph{individual} always retains their spatial location,
so that $\kappa = 0$ corresponds to the parent being assigned a new level,
and our line of descent switching to the offspring.

@@

First, for each $1 \le k < n$
let $(s_k, v_k, y_k, \kappa_k)$ be the $a_{k+1}^\text{th}$ point in $\Pi_{(a)_k}$
(this is the sequence of jumps taken by the line of descent).
Note that $s_k = \tau_{(a)_k}$, and also set $s_0 = 0$.
Let $\Gamma_a$ by the point process ``experienced by the line of descent leading to $a$'',
which is for $A \subset [0,\infty)^2 \times \IR^d \times \{0,1\}$
\comment{This is wrong, because $s$ is relative to birth time.}
\begin{align*}
\Gamma_a(A)
    &=
    \sum_{k=1}^n
    \Pi_{(a)_k}\left(A \cap [s_{k-1}, s_k) \times [0, \infty) \times \IR^d \times \{0,1\}\right) 
    \\ &\qquad {}
    + \sum_{k=1}^{n-1}
      \delta_{(s_k, v_k, y_k, 1 - \kappa_k)}(A) .
\end{align*}
Note that we have included the points at which the line of descent's level jumps,
flipping $\kappa$ at these points,
and as a result $\Gamma_a$ has the same distribution as $\Pi_a$
restricted to $[0, \sigma_a)$.
Then for $0 \le t \le \sigma_a$,
\begin{align*}
    X_a(t)
    &=
    X_{a_1}(0)
    +
    \int_{
        [0, t) \times [0, \infty) \times \IR^d \times \{0, 1\}
    }
    \kappa y
    \ind_{
        U_{(a)_{\zeta(s)}}(s) + \ell( X_{(a)_{\zeta(s)}}(s-), y, \eta(s-), v) < \lambda
    }
    \Gamma_{a}(ds, dv, dy, d\kappa) 
\end{align*}
and
\begin{align*}
    U_a(t)
    &=
    U_{a_1}(0)
    +
    \sum_{k=1}^{\zeta_a(t) - 1}
    \frac{
        v_k
    }{
        2 \lambda^{-1} \theta \gamma(X_{(a)_k}(s_k-), \eta(s_k-))
        r(X_{(a)_k}(s_k-) + y_k, \eta(s_k-))
    }
    \\ & \qquad {}
    +
    \int_0^t
    \bigg(
        \frac{\theta}{\lambda} \gamma(X_a(s),\eta(s))
        \int_{\IR^d} r(z,\eta(s)) q(X_a(s),dz) U_a(s)^2
    \\ & \qquad\qquad  \qquad {}
        -
        b(X_a(s),\eta(s)) U_a(s)
    \bigg)
    dt .
\end{align*}

%% %% %% %% %% %% %% %% %% %% %% %%
\subsection{Limiting processes for lines of descent}
\label{sec:limiting_lines_of_descent}

@@

Supposing that they do,
consider how a line of descent $(X_a(t), U_a(t))$ evolves.
The line of descent throws off a new line of descent at a higher level
when there is a point $(s, v, y, \kappa)$ in $\Pi_a$ with 
\begin{equation} \label{eqn:line_points}
v < 2 (\lambda - U_a(t)) \lambda^{-1} \theta \gamma(X_a(t), \eta) r(X_a(t)+y, \eta) .
\end{equation}
Since the mean measure of the $v$ coordinate is Lebesgue,
$\theta/\lambda \to \alpha$,
and $q(x, dy) \to \delta_x(dy)$,
this corresponds in the limit to new lines of descent being thrown off at all times,
i.e., according to a Poisson process with rate
$$
\alpha \gamma(X_a(t), \eta) r(X_a(t), \eta) 
$$
and uniform intensity on higher levels.
Again, since in the limit the dispersal distance is zero,
new levels are produced at the same spatial location.
However, this affects the location of the line of descent:
for each new line of descent, there is a $1/2$ chance
that the line of descent switches location (i.e., moves to $X_a(t) + y$).
Taking $g$ to be a suitable test function on $\IR^d$, 
and rewriting~\eqref{eqn:line_points},
the generator of this motion is, when the level is $u$,
\begin{align*}
    g(x)
    &\mapsto
    \left(1 - \frac{u}{\lambda}\right) \gamma(x, \eta)
    \theta \int_{\IR^d} r(x+y, \eta) (g(x+y) - g(x)) q(x, dy) \\
    &=
    \left(1 - \frac{u}{\lambda}\right) \gamma(x, \eta)
    \bigg\{
        \theta \int_{\IR^d} (r(x+y, \eta) g(x+y) - r(x, \eta) g(x)) q(x, dy) \\
        &\qquad \qquad {}
        -
        \theta \int_{\IR^d} (r(x+y, \eta) - r(x, \eta) ) g(x) q(x, dy) 
    \bigg\} \\
    &\to
    \gamma(x, \eta)
    \left(
        \DG(rg)(x) - g(x) \DG(r)(x)
    \right) .
\end{align*}
Notice that the factors of 2 have canceled,
and that the result is independent of $u$.

@@

To write out the joint motion of spatial coordinate and levels,
it is useful to work out the differential operator above in more detail.
Recall that $\DG g(x) = \sum_i \meanq_i \partial_i g(x) + \sum_{ij} \covq_{ij} \partial_{ij} g(x)$,
and for the moment write $r(x)$ for $r(\cdot, \eta(\cdot))(x)$,
so that
\begin{align}
\DG(rg)(x) - g(x) \DG(r)(x)
    &= \nonumber
    r(x) \sum_i \meanq_i \partial_i g(x)
    + 2 \sum_{ij} \meanq_i \partial_i r(x) \covq_{ij} \partial_j g(x)
    + r(x) \sum_{ij} \covq_{ij} \partial_{ij} g(x) \\
    &= \label{eqn:limiting_generator}
    r(x) \left\{
        \left(
        \meanq
        + 2 \covq \grad \log r(x)
        \right)
        \cdot
        \grad g(x)
        +
        \sum_{ij} \covq_{ij} \partial_{ij} g(x)
    \right\} .
\end{align}

@@

Now, let $\beta(x, \eta) = \meanq + 2 \Sigma \grad \log r(x, \eta)$,
and associate with each label an independent $d$-dimensional Brownian motion $W_a$,
and let $\covq = K K^T$.
Then, we can summarize the results above by saying that
in the infinite density limit,
the spatial locations and level of each line of descent $a$
behave for $\tau_a \le t < \sigma_a$ as
\begin{align*}
X_a(t)
    &=
    X_a(\tau_a)
    + \int_{\tau_a}^{t}
    \bigg(
        \beta(x, \eta(s)) ds
        +
        K dW_a(s)
    \bigg)
    ds  \\
U_a(t)
    &=
    U_a(\tau_a)
    + \int_{\tau_a}^{t}
    \bigg(
        \alpha \gamma(X_a(s),\eta(s))
        r(x,\eta(s)) U_a(s)^2
\\ &\qquad \qquad \qquad {}   
        -
        \left\{
            \gamma(X_a(s),\eta(s)) \DG r(X_a(s),\eta(s))
            + F(X_a(s), \eta(s))
        \right\}
        U_a(s)
    \bigg)
    ds .
\end{align*}
To complete the description,
for each label $a \in \labelspace$ let $R_a$ denote an independent Poisson process on $[0, \infty)^2$
with Lebesgue mean measure,
and as before order the points of each in some way.
If the population state at time 0 is $\eta(0)$,
then let $(x_i, u_i)$ be the points of a Poisson process
with mean measure on $\IR^d \times [0, \infty)$ a product of $\eta(0)$ and Lebesgue measure.
\comment{Here I've switched $t$ to be the absolute birth time
rather than relative to the birth time of the parent; I think it's clearer?
It, should be made to agree with the previous section.}
For each point begin a line of descent
with label $i$, location $X_i(0) = x_i$, level $U_a(0) = u_i$, and birth time $\tau_i = 0$.
Each point in the $R_a$ denotes a potential birth time for $a$:
if the $j^\text{th}$ point in $R_a$ is $(t, v)$,
$\tau_a \ge t$, and and $U_a(t) < \infty$,
then a new line of descent with label $aj$ is produced,
with birth time $\tau_{aj} = t$, location $X_{aj}(t) = X_a(t)$, and level
$$
    U_{aj}(t) = U_a(t)
    + \frac{v}{ 2 \alpha \gamma(X_a(t), \eta(t)) r(X_a(t), \eta(t)) } .
$$
\comment{Perhaps the conclusion of the Propostion would be something like:}
We can now express our population processes in the limit using these levels:
\begin{align*}
    \eta_t = \sum_{a : \tau_a > t \;\&\; U_a(t) < \lambda} \delta_{X_a(t)}
    \qquad \text{and} \qquad
    \lp = \sum_{a : \tau_a > t \;\&\; U_a(t) < \lambda} \delta_{(X_a(t), U_a(t))} .
\end{align*}

@@

level. Consequently, for $t \geq \tau$
\begin{align}\label{eq: Spatial evolution of individual lineage}
X_{\tau}(\tau+t)=&X_{\tau}(t)\\
&+\int_{(0,t]\times[0,\infty]\times[0,1]\times\{0,1\}}\ind_{U_{\tau}(\tau+s)+ \frac{v}{2\lambda^{-1} \theta \gamma^{\mathfrak{m}}_{\theta}(X_{\tau}(\tau+s-), \eta(\tau+s-))}<\lambda\}}\kappa    H(X_{\tau}(\tau+s-), \eta(\tau+s),\omega) \\
&\qquad \times \xi_{\tau}(ds,dv,dw,d\kappa).
\end{align}
Starting with an initial particle at level $U_{(0,i)}$ , we can identify a potential line of descent by
taking $(s_1 , v_1 , w_1 , \kappa_1 ) \in \xi_{(0,i)}$ , setting $\tau_1=s_1$ , and then recursively taking $(s_k , v_k , w_k , \kappa_k ) \in \xi_{\tau_k }$ and defining $\tau_{k+1} = \tau_{k} + s_{k+1}$. Each time $\tau$ at which a new level appears has an “ancestory” of this type.\\
We now explain the cases where $\kappa_k=0$ and $\kappa_k=1$. When $\kappa_k=0$, the parent retains the level, so $X_{\tau_{k}}$ remains the same at time $\tau_{k+1}$. When $\kappa_k=1$, the offspring gets its parent's level while the parent inherits a new level sampled between $[U_{\tau_k}(\tau_{k+1}-),\lambda]$. As a result, the spatial location $X_{\tau_{k}}$ jumps from the parent's location to the offspring's location at time $\tau_{k+1}$.  Meanwhile, between $[\tau_k, \tau_{k+1}]$ there maybe further birth events generated by $\xi_{\tau_k}$. If $(s,v,w,\kappa) \in \xi_{\tau_k}$, Equation \eqref{eq: Spatial evolution of individual lineage} holds and gives the spatial evolution of $X_{\tau_k}$ for $t \in [\tau_k, \tau_{k+1}]$.\\

For $\tau=\tau_m$ in a recursion like this, let
\begin{equation}\label{eq: Individual Evolution Recursion Terms}
\zeta^{0}_{\tau}(B)=\sum_{k=1}^{m}\delta_{(s_k,v_k,w_k,(1-\kappa_k))}, ~~~ V_{\tau}(t)=\sum_{k=1}^{m}v_k\ind_{[0, \tau_k]}  (t)  
\end{equation}
and define $U_{\tau}(0)=U_{(0,i)}, X_{\tau}(0)=X_{(0,i)}(0)$, and $\zeta_{\tau}$ by
\begin{equation}\label{eq: integrand in spatial tracing}
\zeta_{\tau}(B)=\zeta_{\tau}^{0}(B)+\sum_{k=0}^{m}\int \ind_{B}(\tau_k+s,v,w,\kappa)\ind_{\{\tau_k+s<\tau_{k+1}\}}\xi_{\tau_k}(ds,dv,dw,d\kappa)    
\end{equation}
